Wigglyhedra

### 2024-02-28 draft/wigglyhedra.tex
Added proof of well-definedness of bijections.

\begin{proof}
%We first have to prove that~$\Phi$ and~$\Psi$ are well-defined.
\vincent{todo}
\end{proof}

>>>>

\begin{proof}
  % We first have to prove that~$\Phi$ and~$\Psi$ are well-defined.
  We break the proof into several steps.
  \subsection*{\(\Phi(T)\) is a well-defined wiggly permutation for each wiggly pseudotriangulation \(T\).}
  Let \(\sigma = \Phi(T)\), and for any \(j \in [n]\), let \(t_j\) be the pseudotriangle of \(T\) that has hinge \(j\).
  Suppose that \(2j-1\) appears before \(2j\) in \(\sigma\) for some fixed \(j\).
  This means that the pseudotriangle \(t_j\) has its unlabelled corner to the left of its labelled corners.
  In other words, its hinge opens to the right (see, e.g., the triangle labelled by the \(2\)nd hinge in~\cref{fig:bijection}).

  Consider any label \(i\) that we encounter as we traverse from \(2j-1\) to \(2j\).
  This label belongs to some pseudotriangle \(t_h\), such that \(j\) is either a labelled or an unlabelled corner of \(t_h\).
  Moreover, the corner angle of \(t_h\) is sandwiched within the convex angle at \(j\) formed by \(t_j\).
  If \(j\) is an unlabelled corner of \(t_h\), then \(t_h\) also opens to the right at \(h\).
  If \(j\) is a labelled corner of \(t_h\), then \(t_h\) opens to the left.
  In both cases, it is clear that \(h > j\), and hence \(i > 2j-1\).
  In other words, \(\sigma\) avoids the pattern \(2j-1 \cdots i \cdots 2j\) for \(i > 2j-1\).

  The other case (when \(2j\) appears before \(2j-1\) in \(\sigma\)) is similar.

  \subsection*{\(\Psi(\sigma)\) is a wiggly pseudotriangulation for each wiggly permutation \(\sigma\).}
  By~\cref{prop:wigglyPseudotriangulations}, it suffices to check that for each \(k \neq k'\) in \([2n-1]\), the arcs 
  \(\alpha(\sigma,k)\) and \(\alpha(\sigma,k')\) are distinct and compatible.

  To check distinctness, suppose that \(\alpha(\sigma,k) = (i,j,A,B)\) equals \((\sigma,k') = (i',j',A',B')\) for some \(k \leq k'\).
  Note that \(i = i',j = j'\) are both in \(\sigma([k])\), and also \(\sigma([k]) \subseteq \sigma([k'])\).
  For any \(l \in \sigma([k']) \ssm \sigma([k])\), the conditions on wiggly permutations imply that
  \[i < \left\lceil \frac{l}{2} \right\rceil < j.\]
  Furthermore, from the definitions of \(i\) and \(j\), we see that for any \(l \in \sigma([k']) \ssm \sigma([k])\), we have
  \[2 \left\lceil \frac{l}{2} \right\rceil \in \sigma([k']) \text{ if and only if } 2\left\lceil \frac{l}{2} \right\rceil - 1 \in \sigma([k']).\]
  Now it follows from the fact that \(A' = A\) that there is no element \(l \in \sigma([k']) \setminus \sigma([k])\); that is, \(k = k'\).

  We check compatibility.
  Suppose that for some \(k < k'\), we have 
  \[\alpha(\sigma,k) = (i,j,A,B) \text{ and }\alpha(\sigma,k') = (i',j',A',B').\]
  It is clear from the definition that \(i \neq j'\), as the sets over which we take the max and min to define \(i\) and \(j'\) respectively are disjoint.
  Similarly, we have \(j \neq i'\).
  Therefore the two arcs are pointed.
  To check that the arcs are non-crossing, we must show that the sets
  \[(A \cap B') \cup (\{i,j\} \cap B') \cup (A \cap \{i',j'\}) \text{ and } (A' \cap B) \cup (\{i',j'\} \cap B) \cup (A' \cap \{i,j\})\]
  are both empty.
  We check the first; the second is similar.
  Recall that \(\sigma([k]) \subset \sigma([k'])\).
  For any \(l \in B'\), we have \(2l, 2l-1 \notin \sigma([k'])\), and hence \(2l, 2l-1 \notin \sigma([k])\), which means that \(l \notin A\).
  So \(A \cap B' = \emptyset\).
  Since \(2i - 1, 2j \in \sigma([k]) \subset \sigma([k'])\), we see that \(\{i,j\} \cap B' = \emptyset\).
  Since \(2i', 2j'-1 \notin \sigma([k'])\) hence not in \(\sigma([k])\), we see that \(A \cap \{i',j'\} = \emptyset\).
  \subsection*{The maps \(\Phi\) and \(\Psi\) are inverse bijections that induce the directed graph isomorphism claimed in the statement.}
  \asilata{Todo.}
\item

\end{proof}
